SimBio: a versatile and expandable Python package for modeling systems.

### 2023-11-23 article.qmd
Add Sheriff suggestions

Another weakness is related to the way models are written.
While most tools support standards such as the System Biology Markup Language (SBML) [@huckaSBMLL3V2],
no one is expected to write an SBML file directly.
Tellurium uses a Domain Specific Language (DSL) called Antimony [@smithAntimonyModularModel2009],
which can be translated to SBML.
A DSL allows code to be reused in different programming environments,
but it will not be recognized by Integrated Development Environments (IDEs)
and therefore cannot provide the development experience we have come to expect, such as
syntax highlighting, code completion, refactoring, and static analysis.
This has been recognized by the community,
and an extension for Visual Studio Code has been created as part of a master thesis [@vscode-antimony].
However, maintaining it in the future could be burdensome for the smaller systems biology community.

We developed SimBio,
a Python package for defining and simulating reaction networks,
to address these two issues.
It focuses on providing an ergonomic experience for end-users
by integrating well with IDEs and static analysis tools
through the use of standard modern Python syntax.
It can be extended from Python
since it is coded in pure Python,
which is beneficial for power users.
Its models can be introspected to create other representations,
such as graphs connecting species and/or reactions,
or tables with parameters or equations.
Finally,
it has a modular architecture with a clear separation of concerns into three packages,
making it easier to maintain or to contribute new code,
which is beneficial for developers and maintainers.

>>>>

Another weakness is related to the way models are written.
While most tools support standards such as the System Biology Markup Language (SBML) [@huckaSBMLL3V2],
no one is expected to write an SBML file directly.
Tellurium uses a Domain Specific Language (DSL) called Antimony [@smithAntimonyModularModel2009],
which can be translated to SBML.
A DSL allows code to be reused in different programming environments,
but it will not be recognized by Integrated Development Environments (IDEs)
and therefore cannot provide the development experience we have come to expect, such as
syntax highlighting, code completion, refactoring, and static analysis.
This has been recognized by the community,
and an extension for Visual Studio Code has been created [@vscode-antimony].
However, maintaining it in the future could be burdensome for the smaller systems biology community.
TODO: this is missing for other tools

We developed SimBio,
an open-source Python package for defining and simulating reaction networks,
to address these two issues.
It focuses on providing an ergonomic experience for end-users
by integrating well with IDEs and static analysis tools
through the use of standard modern Python syntax.
It can be extended from Python
since it is coded in pure Python,
which is beneficial for power users.
TODO: high degree of flexibility,  large potential to leverage existing python tools.  For example, ....
Its models can be introspected to create other representations,
such as graphs connecting species and/or reactions,
or tables with parameters or equations.
Finally,
it has a modular architecture with a clear separation of concerns into three packages,
making it easier to maintain or to contribute new code,
which is beneficial for developers and maintainers.
